Revisiting the Relationship between Short-Chain Fatty Acids, the Microbiota, and Colorectal Tumors

### 2018-07-23 submission/manuscript.Rmd
update fig legends and add needed analysis tables

@@

**Figure 3. No change in butyrate producing genes identified between normal, adenoma, and carcinoma individuals.** Imputed gene relative abundance of important butyrate pathway genes using PICRUSt (A). Counts per million (corrected for size and number of contigs in an OPF) for the Butyrate Kinase gene (B). The other butyrate pathway genes from the PICRUSt analysis did not align to any of the OPFs in the metagenome analysis.

@@

**Figure S1. OTU-based Random Forest models of SCFA concentrations.** The train and test results of 100 different 80/20 OTU-based models with 10-fold CV based on higher or lower than the median SCFA concentration using classification Random Forest (A). The top 10 OTUs based on mean decrease in accuracy (MDA) for each model, colored by their lowest taxonomic identification (B).

>>>>

# Load in the group comparison from SCFA models
class_group_rf_data <- read_csv("../data/process/tables/full_model_class_group_results.csv")
reg_group_rf_data <- read_csv("../data/process/tables/full_model_reg_group_results.csv")

@@

**Figure 3. No change in butyrate producing genes identified between normal, adenoma, and carcinoma individuals.** Imputed gene relative abundance of important butyrate pathway genes using PICRUSt (A). Counts per million (corrected for size and number of contigs in an OPF) for the Butyrate Kinase gene (B). The other butyrate pathway genes from the PICRUSt analysis did not align to any of the OPFs in the metagenome analysis.  
**Figure 4. OTU-based regression Random Forest models of SCFA concentrations.** The train and test correlation between actual and predicted values from 100 different 80/20 split OTU-based models with 10-fold CV using regression Random Forest (A). The model accuracy of predicted SCFA concentrations differed between individuals without tumors, patients with adenomas, and patients with carcinomas. Generally, patients with carcinomas had predicted concentrations closest to their actual measured concentration (B). The top 10 OTUs based on mean decrease in accuracy (MDA) for each SCFA model, colored by their lowest taxonomic identification (C).

**Figure 5. Patients with adenomas had the higest number of significant negative correlations between OTU relative abundance and SCFA concentration.** Colors denote the family or lowest taxonomic ID that an OTU classified to. Fewer significant positive correlations were observed overall. Additionally, the differences in the number of significant positive correlations between patients with adenomas versus individuals without tumors (normal) and patients with carcinomas was not as pronounced as the number of significant negative correlations.

@@

**Figure S1. OTU-based classification Random Forest models of high/low SCFA groups based on overall SCFA median concentration.** The train and test results of 100 different 80/20 split OTU-based models with 10-fold CV based on higher or lower than the median SCFA concentration using classification Random Forest (A). The model accuracy of predicted high/low SCFA groups differed between individuals without tumors, patients with adenomas, and patients with carcinomas. Patients with adenomas consistently had the best classification accuracy (B). The top 10 OTUs based on mean decrease in accuracy (MDA) for each SCFA model, colored by their lowest taxonomic identification (C).

**Figure S2. Patients with adenomas had the higest number of significant differences in OTU relative abundance between high/low SCFA groups.** Colors denote the family or lowest taxonomic ID that an OTU classified to. Fewer significant OTUs were observed in individuals without tumors (normal) and patients with carcinomas versus patients with adenomas.

### 2018-07-23 submission/manuscript.tex
update fig legends and add needed analysis tables

\textbf{Figure 3. No change in butyrate producing genes identified
between normal, adenoma, and carcinoma individuals.} Imputed gene
relative abundance of important butyrate pathway genes using PICRUSt
(A). Counts per million (corrected for size and number of contigs in an
OPF) for the Butyrate Kinase gene (B). The other butyrate pathway genes
from the PICRUSt analysis did not align to any of the OPFs in the
metagenome analysis.

@@

\textbf{Figure S1. OTU-based Random Forest models of SCFA
concentrations.} The train and test results of 100 different 80/20
OTU-based models with 10-fold CV based on higher or lower than the
median SCFA concentration using classification Random Forest (A). The
top 10 OTUs based on mean decrease in accuracy (MDA) for each model,
colored by their lowest taxonomic identification (B).

>>>>

\textbf{Figure 3. No change in butyrate producing genes identified
between normal, adenoma, and carcinoma individuals.} Imputed gene
relative abundance of important butyrate pathway genes using PICRUSt
(A). Counts per million (corrected for size and number of contigs in an
OPF) for the Butyrate Kinase gene (B). The other butyrate pathway genes
from the PICRUSt analysis did not align to any of the OPFs in the
metagenome analysis.\\
\textbf{Figure 4. OTU-based regression Random Forest models of SCFA
concentrations.} The train and test correlation between actual and
predicted values from 100 different 80/20 split OTU-based models with
10-fold CV using regression Random Forest (A). The model accuracy of
predicted SCFA concentrations differed between individuals without
tumors, patients with adenomas, and patients with carcinomas. Generally,
patients with carcinomas had predicted concentrations closest to their
actual measured concentration (B). The top 10 OTUs based on mean
decrease in accuracy (MDA) for each SCFA model, colored by their lowest
taxonomic identification (C).

\textbf{Figure 5. Patients with adenomas had the higest number of
significant negative correlations between OTU relative abundance and
SCFA concentration.} Colors denote the family or lowest taxonomic ID
that an OTU classified to. Fewer significant positive correlations were
observed overall. Additionally, the differences in the number of
significant positive correlations between patients with adenomas versus
individuals without tumors (normal) and patients with carcinomas was not
as pronounced as the number of significant negative correlations.

@@

\textbf{Figure S1. OTU-based classification Random Forest models of
high/low SCFA groups based on overall SCFA median concentration.} The
train and test results of 100 different 80/20 split OTU-based models
with 10-fold CV based on higher or lower than the median SCFA
concentration using classification Random Forest (A). The model accuracy
of predicted high/low SCFA groups differed between individuals without
tumors, patients with adenomas, and patients with carcinomas. Patients
with adenomas consistently had the best classification accuracy (B). The
top 10 OTUs based on mean decrease in accuracy (MDA) for each SCFA
model, colored by their lowest taxonomic identification (C).

\textbf{Figure S2. Patients with adenomas had the higest number of
significant differences in OTU relative abundance between high/low SCFA
groups.} Colors denote the family or lowest taxonomic ID that an OTU
classified to. Fewer significant OTUs were observed in individuals
without tumors (normal) and patients with carcinomas versus patients
with adenomas.
